Measuring precise radial velocities and cross-correlation function line-profile variations using a Skew Normal density\thanks{Based on observations collected at the La Silla Parana Observatory, ESO (Chile), with the HARPS spectrograph at the 3.6-m telescope.}

### 2018-11-23 manuscript_v2_final.tex
Add files via upload
final updates: the file is manuscript_v2_final.tex(pdf)

The proposed function defined in Eq.~\eqref{eq:RV:correction} is the result of statistical and astronomical considerations. 
In particular, we considered issues with correlations between the model covariates, which can lead to issues because the matrix needed to calculate the estimates could be singular; this problem is known in statistics as \textit{multicollinearity} and more discussion can be found in \citet{belsley1991}.
In the analysis of real data presented in this work, we never observed a correlation coefficient exceeding $0.66$ between the asymmetry and width parameters and therefore the problem of multicollinearity appeared to be mitigated. 
We also investigated the statistical significant of the interactions between $A$ and the width of the CCF, and $A$ and the asymmetry of the CCF; those two interactions were not statistically significant.

@@

For a given configuration of spots and faculae on the stellar surface, SOAP 2.0 outputs simulated CCFs as a function of rotational phase. 
The code also returns the RV and the FWHM estimated using a Normal model for the CCF, and the BIS SPAN obtained by calculating the bisector of the CCF. 
\sout{SOAP 2.0 gives noiseless CCFs affected by stellar activity, which are used to compare the performances of SN and Normal models of the CCF.}

For the simulations discussed below, a star similar to the Sun was modelled with a solar disc of one solar radius seen equator-on, and a stellar rotational period set to 25.0 days.
The stellar effective temperature is set to 5778 K, and a quadratic limb-darkening relation with linear and quadratic coefficients 0.29 and 0.34, respectively \citep[][]{Oshagh-2013a, Claret-2011}. {\bf The temperature difference of the spot with the photosphere is $\Delta T_{spot} = -663$ K and the temperature difference of the facula depends on the centre-to-limb angle $\theta$,  $\Delta T_{facula} = 250.9 - 407.7\,\cos{\theta} + 190.9\,\cos^2 \theta$ K \citep[][]{Meunier-2010a}}.
In order to make the result of the simulations more comparable to real data obtained with the HARPS spectrograph discussed in Sect.~\ref{sec:4}, the SOAP 2.0 CCFs were generated with a width of 40 \kms\, and considering initial spectra with resolution of R=115'000.
{\bf{ For the simulations presented in this section, we considered a S/N of 100.}}

@@

Fig.~\ref{fig:spot.corr} shows the correlations between the asymmetry parameters and the different estimates of the CCF barycenter (i.e. SN mean RV, SN median RV and N mean RV). {\bf{The correlation between $\gamma$ and SN median RV and the correlation between the BIS SPAN and N mean RV are the strongest ($R=-0.81$) followed by the $\gamma$ - SN mean RV correlation ($R=-0.76$).}} 
The correlations between the width and the CCF barycenter draw circles and no significant correlation is observed. 
{\bf Not like in the facula case, when considering a spot simulated with SAOP 2.0 and a S/N of 100, the SN fit to the CCF does not gives parameters that can probe better stellar activity.
\sout{Similar to the case with a facula, some parameters of the SN gives stronger correlations compared to the Normal parameters.}}

@@

As before, the original RV estimates are corrected using Eq.~\eqref{eq:RV:correction}. 
The results of this correction are displayed in Fig.~\ref{fig:spot.correction} and the statistical tests on the coefficients involved in Eq.~\eqref{eq:RV:correction} are summarized in Table~\ref{table:spot.test}. 
{\bf In the case of a spot, the proposed correction is not able to perform as well as for the facula, and $R^2$ values for the linear combination are between 0.7 and 0.8}. 
{\bf The correction is able to mitigate stellar activity from a N mean RV rms of 6.14\ms down to 3.04\ms. The corresponding values for SN median RV and SN mean RV are 5.85\ms down to 2.74 \ms and 5.27\ms down to 2.74\ms, respectively.}
When comparing the activity correction proposed in this paper with what is commonly used, which means only a linear combination of the width and asymmetry of the CCF, we see that our solution is capable of reducing the RV residual rms by {\bf{6 \%}}.
The Normal or SN parameters involved in Eq.~\eqref{eq:RV:correction} are statistically significant to explain the activity signal as seen in Table~\ref{table:spot.test}, {\bf except the width of the CCF $\beta_3$ and the interaction term $\beta_4$}
%This is not surprising when looking at the circle shape drawn when plotting the width as a function of the RV in Fig.~\ref{fig:spot.corr}.

@@

The analysis of the simulated SOAP 2.0 data in the previous section were helpful in assessing the performance of the proposed methodology in a setting where the ground truth was known. {\bf We found that the parameters derived by the SN were equally correlated or in the case of a facula even more when the data are only affected by stellar activity.} In this section we present an analysis conducted on real observations, in particular, the star Alpha Centauri B, and compare the performance when fitting a CCF using the SN density defined in Sec. \ref{sec:3} with the commonly employed approach based on fitting a Normal density for estimating the RV and retrieving the BIS SPAN for evaluating the asymmetry of the CCF. 
%to retrieve the RV and width of the CCF and calculating the bisector to derive the asymmetry parameter BIS SPAN. 
Four other stars have been analyzed with the proposed method and details can be found in Appendix \ref{appendix}. 
For all the stars considered in the presented work, only CCFs that were derived from spectra that had at least a S/N of 10 at 550 nm were selected.

@@

Because of the advantages of using the proposed SN approach over the commonly employed approach based on the Normal density fit to the CCF and the BIS SPAN or the asymmetry parameters described in \citet{Boisse-2011} and \citet{Figueira-2013}, a SN density model for the CCF may be more useful for detecting stellar activity than the previously proposed parametrizations.
Correlations between $\gamma$ and SN mean RV, and between the width and SN mean RV can be used to probe stellar activity signals in RV data, and SN median RV can be used to estimate RV.
We also proposed a new model to correct the estimated RV data for stellar activity signals, by using the amplitude of the CCF and an interaction term between the estimated asymmetry and the width parameters. Using simulated data from SOAP 2.0, {\bf{this new proposed correction reduces the effect of the stellar activity signal by an additional 14.5 -15 \% and 5.3-5.8 \%}} over the usual model, respectively, for facula and spot. When applying this model on real data,\sout{the improvement is not as substantial, however, we still} we observe that planetary detection limits are improved by a non-negligible 12\%.

>>>>

The proposed function defined in Eq.~\eqref{eq:RV:correction} is the result of statistical and astronomical considerations. 
{\bf{ In particular, we evaluated the correlations between the model covariates, since high correlations can lead to a non-invertible design matrix resulting in invalid parameter estimates. }}This problem is known in statistics as \textit{multicollinearity} and more discussion can be found in \citet{belsley1991}.
In the analysis of real data presented in this work, we never observed a correlation coefficient exceeding $0.66$ between the asymmetry and width parameters and therefore the problem of multicollinearity appeared to be mitigated. 
We also investigated the statistical significant of the interactions between $A$ and the width of the CCF, and $A$ and the asymmetry of the CCF; those two interactions were not statistically significant.

@@

For a given configuration of spots and faculae on the stellar surface, SOAP 2.0 outputs simulated CCFs as a function of rotational phase. 
The code also returns the RV and the FWHM estimated using a Normal model for the CCF, and the BIS SPAN obtained by calculating the bisector of the CCF. 
%\sout{SOAP 2.0 gives noiseless CCFs affected by stellar activity, which are used to compare the performances of SN and Normal models of the CCF.}

For the simulations discussed below, a star similar to the Sun was modelled with a solar disc of one solar radius seen equator-on, and a stellar rotational period set to 25.0 days.
The stellar effective temperature is set to 5778 K, and a quadratic limb-darkening relation with linear and quadratic coefficients 0.29 and 0.34, respectively \citep[][]{Oshagh-2013a, Claret-2011}. {\bf The temperature difference of the spot with the photosphere is $\Delta T_{spot} = -663$ K and the temperature difference of the facula depends on the centre-to-limb angle $\theta$,  $\Delta T_{facula} = 250.9 - 407.7\,\cos{\theta} + 190.9\,\cos^2 \theta$ K \citep[][]{Meunier-2010a}}.
In order to make the result of the simulations more comparable to real data obtained with the HARPS spectrograph discussed in Sect.~\ref{sec:4}, the SOAP 2.0 CCFs were generated with a width of 40 \kms\, and considering initial spectra with resolution of Res =115'000.
{\bf{ For the simulations presented in this section, we considered a S/N of 100.}}

@@

Fig.~\ref{fig:spot.corr} shows the correlations between the asymmetry parameters and the different estimates of the CCF barycenter (i.e. SN mean RV, SN median RV and N mean RV). {\bf{The correlation between $\gamma$ and SN median RV and the correlation between the BIS SPAN and N mean RV are the strongest ($R=-0.81$) followed by the $\gamma$ - SN mean RV correlation ($R=-0.76$).}} 
The correlations between the width and the CCF barycenter draw circles and no significant correlation is observed. 
{\bf Unlike the facula scenario, when considering a spot simulated from SOAP 2.0 and a S/N of 100 the SN parameters do not appear to better probe stellar activity than the Normal parameters.}
%\sout{Similar to the case with a facula, some parameters of the SN gives stronger correlations compared to the Normal parameters.}}

@@

As before, the original RV estimates are corrected using Eq.~\eqref{eq:RV:correction}. 
The results of this correction are displayed in Fig.~\ref{fig:spot.correction} and the statistical tests on the coefficients involved in Eq.~\eqref{eq:RV:correction} are summarized in Table~\ref{table:spot.test}. 
{\bf In the case of a spot, the proposed correction is not able to perform as well as for the facula, and $R^2$ values for the linear combination are between 0.7 and 0.8}. 
{\bf The correction is able to mitigate stellar activity from a N mean RV rms of 6.14\ms down to 3.04\ms. The corresponding values for SN median RV and SN mean RV are 5.85\ms down to 2.74 \ms and 5.27\ms down to 2.74\ms, respectively.}
When comparing the activity correction proposed in this paper with what is commonly used, which means only a linear combination of the width and asymmetry of the CCF, we see that our solution is capable of reducing the RV residual rms by {\bf{5.3 - 5.8 \%}}.
The Normal or SN parameters involved in Eq.~\eqref{eq:RV:correction} are statistically significant to explain the activity signal as seen in Table~\ref{table:spot.test}, {\bf except the width of the CCF $\beta_3$ and the interaction term $\beta_4$.}
%This is not surprising when looking at the circle shape drawn when plotting the width as a function of the RV in Fig.~\ref{fig:spot.corr}.

@@

The analysis of the simulated SOAP 2.0 data in the previous section were helpful in assessing the performance of the proposed methodology in a setting where the ground truth was known. {\bf We found with the simulated data that the parameters derived by the SN had correlations comparable to those derived by the Normal; in the case of a facula, the SN parameters had higher correlations for the FWHM than those of the Normal.} In this section we present an analysis conducted on real observations, in particular, the star Alpha Centauri B, and compare the performance when fitting a CCF using the SN density defined in Sec. \ref{sec:3} with the commonly employed approach based on fitting a Normal density for estimating the RV and retrieving the BIS SPAN for evaluating the asymmetry of the CCF. 
%to retrieve the RV and width of the CCF and calculating the bisector to derive the asymmetry parameter BIS SPAN. 
Four other stars have been analyzed with the proposed method and details can be found in Appendix \ref{appendix}. 
For all the stars considered in the presented work, only CCFs that were derived from spectra that had at least a S/N of 10 at 550 nm were selected.

@@

Because of the advantages of using the proposed SN approach over the commonly employed approach based on the Normal density fit to the CCF and the BIS SPAN or the asymmetry parameters described in \citet{Boisse-2011} and \citet{Figueira-2013}, a SN density model for the CCF may be more useful for detecting stellar activity than the previously proposed parametrizations.
Correlations between $\gamma$ and SN mean RV, and between the width and SN mean RV can be used to probe stellar activity signals in RV data, and SN median RV can be used to estimate RV.
We also proposed a new model to correct the estimated RV data for stellar activity signals, by using the amplitude of the CCF and an interaction term between the estimated asymmetry and the width parameters. Using simulated data from SOAP 2.0, {\bf{this new proposed correction reduces the effect of the stellar activity signal by an additional 14.5 -15 \% and 5.3-5.8 \%}} over the usual model, respectively, for facula and spot. When applying this model on real data, we observe that planetary detection limits are improved by a non-negligible 12\%.
